---
title: Modules in R
subtitle: eRum 2016
author: Sebastian Warnholz
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Motivation

  - `dplyr` and `data.table`
  - syntax is a plus
      - not with package development
      - rewrite everything after running `R CMD check`
  - performance
  - integration with custom class definitions
      - not supported
  - improve and integrate the definition of anonymous functions (see for example `rlist` and recently `purrr`)


## The 'Solution'

  - extensive use of the formula class
  - pro: 
      - `R CMD check` friendly
      - not relying on non-standard-evaluation: normal dispatch and argument
      passing possible
  - contra:
      - using formulas is a matter of preference


## Aim of the Package

  - don't loose the performance improvements from `dplyr` or `data.table`
  - reuse as much as possible
  - be generic
  - respect custom classes
  - abstract (my) typical (data manipulation) patterns


## Design Issues

  - `dplyr`: functions do one thing only!
      - `slice` and `filter`
  - `data.table`: let's write one function only!
      - `[.data.table` has about 1500 lines of code
      - 400 open issues on github
  - `dat`: I don't believe in extremes
      - is somewhere in between
      - preserves the class (S3 + S4) and uses (mostly) dplyr in the background


## Examples: rows

```{r echo=TRUE}
data("airquality")
suppressPackageStartupMessages(library("dat"))
airQuality <- as.DataFrame(airquality)
airQuality[~Day < 7, ]
```


## Examples: rows

```{r echo=TRUE}
suppressPackageStartupMessages(library("data.table"))
as.data.table(airquality) %>% mutar(~ Day < 7)
```


## Examples: cols

```{r eval = FALSE, echo = TRUE}
airQuality[is.numeric]
# airquality(sapply(airquality, is.numeric))

as.data.table(airquality) %>% mutar("Ozone:Wind") # remeber dplyr::select

mutar(airquality, "^Wi") # regex is triggered by '^'
airQuality["^Wi"]
# airquality[grepl("^Wi", names(airquality))]

airQuality[1] # error - no positional matching!

# Yes, base R syntax is otherwise okay:
ind <- 1:5
airQuality[ind, "Wind"]
```


## Examples: mutate + summary

```{r echo = TRUE, eval = FALSE}
airquality %>%
  mutar(
    meanWind ~ mean(Wind), 
    by = "Month"
  )

airQuality[meanWind ~ mean(Wind), by = "Month"]
```

- `mutar` can do
    - row subsetting
    - col subsetting (character + logical, regex, functions)
    - create new columns using `dplyr::mutate`
    - data aggregation using `dplyr::summarise` (triggered with `by`)
    - stops here


## Integration of Custom Classes

```{r error=TRUE, echo=TRUE}
setClass("MyData", contains = "data.table")
MyData <- function(...) new("MyData", data.table(...))
dat <- do.call(MyData, airquality)
dat[1, ] # it is just broken - same for dplyr
```


## `mutar` can do more!

```{r echo=TRUE}
setMethod("[", "MyData", mutar)
dat[1, ]
dat[1, CopyOfDay ~ Day]
```


## Additional functions

- `map`: as a replacement for `lapply` and `mapply`
- `flatmap`: instead of `sapply`
- `replace`: to replace values in a vector
- `extract`: to extract values from a vector


## Be `data.frame` aware

```{r echo = TRUE}
ind <- sapply(airquality, is.integer)
airquality[ind] <- lapply(airquality[ind], function(x) as.character(x + 1))
map(airquality, x ~ as.integer(x) - 1L, is.character) %>%
  str
```


## Be `list` aware

```{r}
l <- list(
  a = list(name = "Me", age = 29),
  b = list(name = "You", age = ">30")
)
```

```{r echo=TRUE, eval=FALSE}
l <- list(
  a = list(name = "Me", age = 29),
  b = list(name = "You", age = ">30")
)

lapply(l, function(x) x["name"])
map(l, "name")
flatmap(l, "name")
```

```{r echo=TRUE, eval = TRUE}
flatmap(l, ~ .$name) # function(.) .$name
```


## Be paranoid  

- add runtime type checks

```{r echo=TRUE, error=TRUE}
flatmap(l, numeric : x ~ x$age)
flatmap(l, character(1) : x ~ x$name)
```


## Be comprehensive

```{r echo = TRUE}
extract(1:20, ~ . %% 2 == 0)
replace(1:10, ~ . %% 2 == 0, NA_real_)

vec <- 1:20
vec[vec %% 2 == 0]
```


## Summary

- you can fall back to or integrate `dplyr` and `data.table` functions
- syntactic sugar for anonymous functions
- use custom data types:
    - everything is a method: `map`, `replace`, and `extract`
    - `mutar` respects the data it operates on (for better or worse)
- have *pipeable* syntax: no replacement functions
- may not replace everything in an interactive data analysis `->` it covers
almost all use cases for production code


## Want more?

- Modules is on GitHub
    - [github.com/wahani/modules](https://www.github.com/wahani/modules)
    - Just file an issue / make a pull request
    - It has a vignette / Readme

- The Presentation is also in GitHub
    - [github.com/wahani/presiModules](https://www.github.com/wahani/presiModules)

- Sebastian Warnholz
    - wahani@gmail.com
